---
title: "Tank VOC Emissions"
author: "Jon Wilkey"
date: "June 1, 2016"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
    number_sections: true
    fig_width: 6.5
    fig_height: 6.5
    fig_caption: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# 1.1 Paths ---------------------------------------------------------------

# Predefine list object "path" for holding directory path listings
path <- NULL

# Path switch - replace with the path directory for your local copy of the Git
# repository and Dropbox files.
pwd.drop <- "C:/Users/jonwi/"
pwd.git  <- "C:/Users/jonwi/Documents/R/"
  
# Define paths.
# "raw"  is raw data (*.dbf files from DOGM, *.csv files, etc.). 
# "data" is prepared data files (typically *.rda).
# "look" is lookup tables. 
# "plot" is the directory for saving plot *.pdf files.
# "work" is the working directory where main.R and IO_options.R are located.
# "fun"  is the directory for all *.R functions.
path$raw  <-    paste(pwd.drop, "Dropbox/UBOG/Raw Data", sep = "")
path$data <-    paste(pwd.drop, "Dropbox/UBOG/Prepared Data", sep = "")
path$look <-    paste(pwd.drop, "Dropbox/UBOG/Lookup Tables", sep = "")
path$plot <-    paste(pwd.drop, "Dropbox/UBOG/Plots", sep = "")
path$work <-    paste(pwd.git,  "ub_o-g_emissions", sep = "")
path$fun  <-    paste(pwd.git,  "ub_o-g_emissions/Functions", sep = "")
path$plotfun <- paste(pwd.git,  "ub_o-g_emissions/Functions/Plotting", sep = "")

# Remove temporary path objects
remove(pwd.drop, pwd.git)

# Set working directory
setwd(path$work)


# 1.2 Functions -----------------------------------------------------------

# List of functions used in this script to be loaded here
flst <- file.path(path$fun, c("GBMsim.R",
                              "EIAsim.R",
                              "EIAsimLT.R",
                              "drillsim.R",
                              "priorProd.R",
                              "priorInfo.R",
                              "sim_tdelay.R",
                              "sim_DCC.R",
                              "sim_depth.R",
                              "sim_EF.R",
                              "sim_fieldnum.R",
                              "sim_county.R",
                              "sim_lease.R",
                              "sim_rework.R",
                              "sim_dupRework.R",
                              "sim_tdrill.R",
                              "sim_wellType.R",
                              "productionsim.R",
                              "priorProdReworkAdjust.R",
                              "Ecalc.R",
                              "LOCcalc.R",
                              "calc_E_wc.R",
                              "calc_E_rt.R",
                              "calc_E_sh.R",
                              "calc_E_dh.R",
                              "calc_E_tank.R",
                              "calc_E_truck.R",
                              "calc_E_pctrl.R",
                              "calc_E_ppump.R",
                              "calc_E_fug.R",
                              "clipboard.R",
                              "inf_adj.R",
                              "CDFd.R",
                              "CDFq.R",
                              "asYear.R",
                              "NAoverwrite.R"))

# Load each function in list then remove temporary file list variables
for (f in flst) source(f); remove(f, flst)


# 1.3 Packages -----------------------------------------------------------

library(foreign)
library(plyr)
library(zoo)
library(data.table)
library(sqldf)
library(minpack.lm)
library(scatterplot3d)
library(beepr)
library(fitdistrplus)
library(lhs)
library(xtable)
library(knitr)


# 1.4 Options -------------------------------------------------------------

# Don't want strings 'typed' as factors but as characters
options(stringsAsFactors=FALSE)

# Run scripts to load user defined input/output options
source("IO_options.R")
source("EF_options.R")

# Set seed for random number generation (for reproducibility)
set.seed(1)

# 1.5 Load Data -----------------------------------------------------------

# Load data sets used in analysis below
load(file.path(path$data, paste("emissionUpdate_", opt$file_ver, ".rda", sep = "")))

# File path for OGEI *.csv tables
pOGEI <- file.path(path$raw, "OGEI")

# Read in each OGEI table
apis <-        read.csv(file.path(pOGEI, "apis.csv"))                  # API #'s associated with each facility ID
comp.count <-  read.csv(file.path(pOGEI, "component_count.csv"))       # Component counts
dehy <-        read.csv(file.path(pOGEI, "dehydrators.csv"))           # Dehydrators
fac.list <-    read.csv(file.path(pOGEI, "facilities_list.csv"))       # Facilities list
fugitives <-   read.csv(file.path(pOGEI, "fugitives.csv"))             # Fugitive emissions
oper.info <-   read.csv(file.path(pOGEI, "operator_info.csv"))         # Operator information
pneum.ctrl <-  read.csv(file.path(pOGEI, "pneumatic_controllers.csv")) # Pneumatic controllers
pneum.pumps <- read.csv(file.path(pOGEI, "pneumatic_pumps.csv"))       # Pneumatic pumps
prod.areas <-  read.csv(file.path(pOGEI, "production_areas.csv"))      # Production areas (keep?)
rice.EF <-     read.csv(file.path(pOGEI, "rice_emission_factors.csv")) # RICE emission factors
rice.turb <-   read.csv(file.path(pOGEI, "rice_turbines.csv"))         # RICE turbines
sep.heat <-    read.csv(file.path(pOGEI, "separators_heaters.csv"))    # Separators and heaters
tanks <-       read.csv(file.path(pOGEI, "tanks.csv"))                 # Tanks
truck <-       read.csv(file.path(pOGEI, "truck_loading.csv"))         # Truck loading
well.compl <-  read.csv(file.path(pOGEI, "well_completions.csv"))      # Well completions
```

# Introduction

During the training webinar held on May 18, 2016, Whitney and I discussed options for estimating VOC emissions from tanks in the R model. Currently, the model randomly selects total VOC emissions for a tank from a cumulative probability table (CPT) based on data from the "tanks" table of the OGEI database. For example, suppose that a tank had total VOC emissions of 1.0 ton/yr, and that there was only 1 well associated with that tank. If a well randomly selected the row in the CPT for that particular tank, then the tank VOC emissions for that well would always be 1 ton/yr, or in the time steps used in the model, 0.083 ton/month. That emissions rate would be constant, and wouldn't vary with production rate.

The reason that I adopted this approach was that during the data analysis phase I found a strange relationship between VOC emissions and oil throughput:

```{r oil.no.na}
plot(tanks$throughput_oil, tanks$total_voc,
     xlab = "Annual Oil Throughput (bbl / yr)",
     ylab = "Total VOC Emissions (tons / yr)",
     main = "VOC Emissions vs. Oil Throughput")
mtext("Note: Excluding records with NA values for throughput_oil in tanks table")

# For future analysis below
temp <- subset(tanks, subset = (throughput_oil > 0), select = c("total_voc", "throughput_oil"))
temp$r <- temp$total_voc/temp$throughput_oil
```

During my first implementation attempt for tank VOC emissions, I calculated the ratio $r$ of tons VOCs emitted per barrel of oil throughput:

$$\begin{equation}
  r = \dfrac{VOC}{oil}
\end{equation}$$

and then randomly selected from the the $r$ ratio in the CPT. The total VOC tank emissions are then:

$$\begin{equation}
  VOC = r * oil
\end{equation}$$

where $oil$ is the vector of monthly oil production for a given well. The problem that I ran into is that $r$ spans a huge range, from a minimum of ~1e-07 to a maximum of ~22. A CDF for $r$ is shown below (neglecting those values of $r$ that are $\infty$ due to dividing by zero):

```{r rCDF}
plot(CDFq(vector = temp$r, xq = opt$xq),
     xlim = c(1e-7, 1e1),
     log = "x",
     type = "l",
     xlab = "r (log-scale)",
     ylab = "Cumulative Probability",
     main = "CDF of r Ratio")
```

So if you randomly pick $r$ from a CPT, you're fairly likely to combine it with a mismatched oil production vector, producing either too little or too much VOCs.

Given the importance of tank VOCs in the overall emissions inventory, Whitney suggested that we spend more time on trying to refine the approach used so that we can still model emissions as a function of tank throughput (or some other factor that is related to the well(s) feeding into the tank). Plotted below are a series of exploratory plots that look at how VOC emissions vary as a function of fluid throughput, by:

* Fluid type (oil, condensate, and water)
* Control Type (controlled vs. uncontrolled)
* Operator

To cut down on the total number of plots, only the ten largest operators are considered, which are:

```{r opers}
top.op <- sqldf("select operator_id, count(*) as count from tanks group by operator_id order by count desc")[1:10,]
print(top.op, row.names = F)
```

Note - count is the number of records in the tanks table for each operator ID number.

# Exploratory Plots

Notes:

* In all plots black points represent uncontrolled sources, while red points indicate emissions from tanks with controls.
* Any NA values in the data set for fluild throughput volume are assumed to be = 0
    * `r length(which(is.na(tanks$throughput_oil)))` records for oil and `r length(which(is.na(tanks$throughput_condensate)))` records for condensate are NA out of `r nrow(tanks)` total records in the tanks table.
    * Interestingly, only `r length(which(is.na(tanks$throughput_water)))` records for water are NA).
* NA values for control type are assumed to be equivalent to "No Control".

## All Operators

```{r all.op}
# Create dataset for future operations
fld <- data.frame(voc = NA.overwrite(tanks$total_voc),
                  oil = NA.overwrite(tanks$throughput_oil),
                  con = NA.overwrite(tanks$throughput_condensate),
                  wat = NA.overwrite(tanks$throughput_water),
                  ctrl = ifelse(test = (tanks$control_type == "Combustor" |
                                        tanks$control_type == "VRU"),
                                yes  = T,
                                no   = F),
                  oper = tanks$operator_id)

# Oil
plot(voc ~ oil, fld[which(!fld$ctrl),],
     xlim = c(0, 1.1*max(fld$oil)), ylim = c(0, 1.1*max(fld$voc)),
     xlab = "Oil (bbl/yr)", ylab = "VOC (ton/yr)", main = "Oil")
points(voc ~ oil, fld[which(fld$ctrl),], col = "red")

# Oil < 99th percentile
plot(voc ~ oil, fld[which(!fld$ctrl & fld$oil <= quantile(fld$oil, 0.99)),],
     xlab = "Oil (bbl/yr)", ylab = "VOC (ton/yr)", main = "Oil - up to 99th %")
points(voc ~oil, fld[which(fld$ctrl & fld$oil <= quantile(fld$oil, 0.99)),], col = "red")

# Condensate
plot(voc ~ con, fld[which(!fld$ctrl),],
     xlim = c(0, 1.1*max(fld$con)), ylim = c(0, 1.1*max(fld$voc)),
     xlab = "Condensate (bbl/yr)", ylab = "VOC (ton/yr)", main = "Condensate")
points(voc ~con, fld[which(fld$ctrl),], col = "red")

# Condensate < 99th percentile
plot(voc ~ con, fld[which(!fld$ctrl & fld$con <= quantile(fld$con, 0.99)),],
     xlab = "Condensate (bbl/yr)", ylab = "VOC (ton/yr)", main = "Condensate - up to 99th %")
points(voc ~con, fld[which(fld$ctrl & fld$con <= quantile(fld$con, 0.99)),], col = "red")

# Water
plot(voc ~ wat, fld[which(!fld$ctrl),],
     xlim = c(0, 1.1*max(fld$wat)), ylim = c(0, 1.1*max(fld$voc)),
     xlab = "Water (bbl/yr)", ylab = "VOC (ton/yr)", main = "Water")
points(voc ~ wat, fld[which(fld$ctrl),], col = "red")

# Water <= 99th percentile
plot(voc ~ wat, fld[which(!fld$ctrl & fld$wat <= quantile(fld$wat, 0.99)),],
     xlab = "Water (bbl/yr)", ylab = "VOC (ton/yr)", main = "Water - up to 99th %")
points(voc ~ wat, fld[which(fld$ctrl & fld$wat <= quantile(fld$wat, 0.99)),], col = "red")
```

## Individual Operators

### Operator ID # `r top.op$operator_id[1]`

```{r op1}

# Functionalize to save space and time
operplots <- function(op_id) {
  
  # Subset to just specific operator
  temp <- fld[fld$oper == top.op$operator_id[op_id],]
  
  # Oil
  plot(voc ~ oil, temp[which(!temp$ctrl),],
       xlim = c(0, 1.1*max(temp$oil)), ylim = c(0, 1.1*max(temp$voc)),
       xlab = "Oil (bbl/yr)", ylab = "VOC (ton/yr)", main = "Oil")
  points(voc ~ oil, temp[which(temp$ctrl),], col = "red")
  
  # Condensate
  plot(voc ~ con, temp[which(!temp$ctrl),],
       xlim = c(0, 1.1*max(temp$con)), ylim = c(0, 1.1*max(temp$voc)),
       xlab = "Condensate (bbl/yr)", ylab = "VOC (ton/yr)", main = "Condensate")
  points(voc ~con, temp[which(temp$ctrl),], col = "red")
  
  # Water
  plot(voc ~ wat, temp[which(!temp$ctrl),],
       xlim = c(0, 1.1*max(temp$wat)), ylim = c(0, 1.1*max(temp$voc)),
       xlab = "Water (bbl/yr)", ylab = "VOC (ton/yr)", main = "Water")
  points(voc ~ wat, temp[which(temp$ctrl),], col = "red")
}

# Call plot function
operplots(op_id = 1)
```

### Operator ID # `r top.op$operator_id[2]`
```{r op2}
operplots(op_id = 2)
```

### Operator ID # `r top.op$operator_id[3]`
```{r op3}
operplots(op_id = 3)
```

### Operator ID # `r top.op$operator_id[4]`
```{r op4}
operplots(op_id = 4)
```

### Operator ID # `r top.op$operator_id[5]`
```{r op5}
operplots(op_id = 5)
```

### Operator ID # `r top.op$operator_id[6]`
```{r op6}
operplots(op_id = 6)
```

### Operator ID # `r top.op$operator_id[7]`
```{r op7}
operplots(op_id = 7)
```

### Operator ID # `r top.op$operator_id[8]`
```{r op8}
operplots(op_id = 8)
```

### Operator ID # `r top.op$operator_id[9]`
```{r op9}
operplots(op_id = 9)
```

### Operator ID # `r top.op$operator_id[10]`
```{r op10}
operplots(op_id = 10)
```

# Conclusions

Fromt the plots shown here, I think we can draw the following conclusions:

* There's no real relationship between water throughput and VOC emissions.
* There's clearly a linear relationship between oil and/or condensate throughput and VOC emissions, but it varies by operator.
* Most operators are reporting either all oil or all condensate moving through their tank batteries.
    * Oil only operators: 23, 22, 8, 25, 31
    * Condensate only operators: 11, 14, 19
    * Operators with both oil and condensate: 12, 1
* In general, several plots show an inverse relationship between emissions and throughput (i.e. high VOC emissions with low oil/condensate throughput, and vice-versa).
    * This could just be because of new wells, which presumably have the highest production rates and are feeding into new tank batteries with the best controls.
* Other factors besides those contained in tanks table are likely influencing emission rates (tank age, jurisdiction, etc.). I could try including factors located in other tables (anything in the OGEI database or any of the well information that we have from UDOGM), but it would take additional time to program the required merges. If there are any other factors that you would like me to investigate, please let me know.
